---
title: "lionfishanalysis_kb_v3"
author: "Kim Bourne"
date: "May 15, 2018"
output: pdf_document
fontsize: 12pt
---

```{r,echo = FALSE, results = "hide", include = FALSE}
setwd('C:/Users/Kimberly Bourne/Box Sync/Lionfish-Amelia')
data_all <- read.csv('lionfishdata.csv')

data_x <-data.frame(Starve_time = data_all$Starvationtime_hours,
                    cloudcover = data_all$Cloudcover,sex = data_all$Sex,
                    lf_std_length = data_all$Lionfishstandardlength_cm, lf_wetwgt = data_all$LionfishWetWeight_g,
                    mooncycle = data_all$Mooncycle,depthcaptured = data_all$Depthcaptured_m,bodycond = data_all$BodyCondition,Lionfish_totlength = data_all$Lionfishtotallength_cm)
counts <- data.frame(N_chromis = data_all$Numberchromisconsumed, N_goby = data_all$Numbergobyconsumed,N_wrasse = data_all$Numberwrasseconsumed)
```

Using zCompositions, I replaced the zero proportions for the count data. Since sqrt(n) > D, the SQ method was used for the bayesian-multiplicative treatment.

```{r}
# Use zCompositions to replace zeros using bayesian multiplicative methods
library('zCompositions')
c_tot <- counts$N_chromis + counts$N_wrasse + counts$N_goby

p <- cbind(counts$N_chromis/c_tot,counts$N_wrasse/c_tot,counts$N_goby/c_tot) # proportions of each species based on total eaten
rep_p <- cmultRepl(p,method = "SQ",output = "counts") # proportions data with zeros replaced
colnames(p) <- c("P_chromis","P_wrasse","P_goby")
props_sum <- numeric()
for (i in 1:20){props_sum <- c(props_sum, sum(p[i,]))}
props_sum
```

I inserted my own ilr transform function at this point, because I could not find anything on the transformation matrix that the built in function used. I used the Helmert matrix because I found the interpretation to be intuitive.

```{r}
library('compositions')
## ilr transform of proportions
rep_p_m <- as.matrix(rep_p)
ilr <- function(x, p=0) {
  y <- log(x/geometricmeanRow(x))
  k <- dim(y)[2]
  H <- contr.helmert(k)                       # Dimensions k by k-1
  H <- t(H) / sqrt((2:k)*(2:k-1))             # Dimensions k-1 by k
  return(y %*% t(H))                          # Rotated/reflected values
}
p_ilr <- ilr(rep_p_m)
colnames(p_ilr) <- c("C_W","CW_G")
```
